---
title: 'Notebook for Project 01'
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
    number_sections: true
my-var: "monte"  # https://bookdown.org/yihui/rmarkdown/html-document.html
---

You will need to populate this notebook based on the idea of "zero to hero" discussed as Data Provenance.

"exploratory data analysis" We try to visualize the data and articulate some initial findings.
```{r}
#read in the data

```

```{r}
#clean up the data
```

```{r}
#plot histogram. #Use function normalDiagnosticPlot(data) from 06_correlation_notebook_instructor



```

```{r}
#possible ideas

#Think about whether samples dependent or independent? I think samples are independent, but measurements within a sample would be dependent.  But maybe since there are family members, this isn't true?

#Compare means of two independent samples (male versus female?) #t.test(data1, data2, paired=FALSE, var.equal=FALSE); # Are we assuming equal or unequal variance? Use a different method depending on each.

```


```{r}
#review the correlations of various components of the body. 

#Correlation is a technique for investigating the relationship between two quantitative, continuous variables, for example, age and blood pressure. Pearson's correlation coefficient (r) is a measure of the strength of the association between the two variables.

#cor.test(x.0.1, y.0.1.max, method="pearson");  # default

#plot correlation
#corrplot::corrplot( (cor(iris.df)) ); 

```

```{r}
#Create a dendrogram/ evaluate clustering?
#dist(x)  Creates a distance function
#hclust creates a dendrogram.
#consider whether the data should be scaled scale(data)
#library(pvclust);  # install.packages("pvclust", dependencies=TRUE); #draws boxes around the clusters.  Gives fewer clusters than hclust
#pvclust uses "bootstrapping" to analyze the stability of the hclust.  For some reason, you have to transpose the data inputs.
```

You need to formulate some initial Research Questions that would interest you.  The power of "one" suggests there is an overall research question that may have sub-questions.

- e.g., Body Proportions and Artistic Renderings
- e.g., NBA Basketball Players vs Our Sample
- e.g., Studies of Covariates on various aspects of the body measurements

Notice the above are three example topics to be discovered, not an actual research question with sub-questions.  Given the data (and possibly external auxiliary data), what would you enjoy researching?  What would you find intriguing?



[+5 EASTER]  Using RGL or some other 3-D plotting tool, build a stick-figure based on your individual body measurements.  We assume the x-axis starts at the center of the eyes and moves laterally: looking right to the end of your outstretched fingertips is one direction; looking left to the end of your outstretched fingertips in another direction.  The y-axis would be the vertical axis from your feet to your head.  The z-axis would be forward-backward with zero being directly between your two feet (placed 'arm-pit' width apart), half of your shoe size being in the positive direction, the other half in the negative direction.  We could imaging z=0 running directly through your body (kinda like your spine and through your head).  Use line segments to connect parts, and a large point (or circle) where segments connect.  Use a rectangularish-shape for the head if a 3-D ellipsoid is too difficult).

https://en.wikipedia.org/wiki/Anatomical_terms_of_location

- x: lateral (right/left)
- y: superior/inferior
- z: anterior/posterior



