---
title: 'Notebook for Project 01'
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
    number_sections: true
my-var: "monte"  # https://bookdown.org/yihui/rmarkdown/html-document.html
---

You will need to populate this notebook based on the idea of "zero to hero" discussed as Data Provenance.

"exploratory data analysis" We try to visualize the data and articulate some initial findings.
```{r deprecated}
#read in the original data
library(devtools);
path.github = "https://raw.githubusercontent.com/MonteShaffer/humanVerseWSU/master/";
source_url ( paste0( path.github, "humanVerseWSU/R/functions-file.R" ) );

path.to.secret = "C:\\Users\\jsmit\\Desktop\\WSU\\DataAnalytics\\STAT419\\_SECRET_\\";

measure1 = utils::read.csv( paste0(path.to.secret, "measure-students.txt"), header=TRUE, quote="", sep="|");

measure1;
unique(measure1$units)

```

```{r read_data}

#read in the cleansed data
library(devtools);
path.github = "https://raw.githubusercontent.com/MonteShaffer/humanVerseWSU/master/";
source_url ( paste0( path.github, "humanVerseWSU/R/functions-file.R" ) );

path.to.secret = "C:\\Users\\jsmit\\Desktop\\WSU\\DataAnalytics\\STAT419\\_SECRET_\\";

measure_data = utils::read.csv( paste0(path.to.secret, "final.measure.txt"), header=TRUE, quote="", sep="|");


#measure_cleaned = ConvertCmToInches(measure)
measure_cleaned = cleanUpData(measure_data)
summary(measure_cleaned);
```


```{r}
#possible ideas

#Think about whether samples dependent or independent? I think samples are independent, but measurements within a sample would be dependent.  But maybe since there are family members, this isn't true?

#Compare means of two independent samples (male versus female?) #t.test(data1, data2, paired=FALSE, var.equal=FALSE); # Are we assuming equal or unequal variance? Use a different method depending on each.

#Analyze symmetry  Do men and women have different symmetry? DOes symmetry change with age? What factors predict dominant hand?  

#Hypothesis: dominant eye, dominant swing, and dominant writing hand are all the same, and are independent of ethnicity, gender, and age.

#is there a relationship between gender and brown eyes versus blue. Head heights for men versus women?

#Relationship between head height to body height 1:8 for both men and women.
#https://illustrationage.files.wordpress.com/2013/04/andrew-loomis-figure-drawing-for-all-its-worth.pdf

# The one to 8 ratio is for adults, so remove samples wih age <18


#Break into 2 dataframs, 1 man one woman
adults = measure_cleaned[measure_cleaned$age >= 18, ]
#male = measure_cleaned[measure_cleaned$gender=='male',]

summary(adults)

mytable2 <- table(adults$gender)
lbls <- paste(names(mytable2), ": ", mytable2, sep="")
pie(mytable2, labels = lbls,  main="Respondent Gender")
#normalDiagnosticPlot(adults$age)
summary(adults$age)

```

```{r}
mytable <- round(prop.table(table(adults$ethnicity))*100,1)
lbls <- paste(names(mytable), ": ", mytable, sep="")
pie(mytable, labels = lbls,   main="Respondent Ethicity")

```



```{r}
#normalDiagnosticPlot(adults$height)
#normalDiagnosticPlot(adults$head.height)

summary(adults$height[adults$gender=="female"])

sub = adults[, c("height", "head.height", "gender")]
sub$eighth = sub$height/8 
sub$percent_diff = ((sub$head.height - sub$eighth)/sub$head.height) * 100
sub

hist(sub$percent_diff[sub$gender=="female"])
hist(sub$percent_diff[sub$gender=="male"])
hist(sub$percent_diff)
sub_f = sub[sub$gender=="female", ]
sub_m = sub[sub$gender=="male", ]
summary(sub_f)
summary(sub_m)
summary(sub$percent_diff)
#normalDiagnosticPlot(sub$percent_diff)

sd(sub$percent_diff, na.rm = TRUE)
sd(sub$percent_diff[sub$gender=="male"], na.rm = TRUE)
sd(sub$percent_diff[sub$gender=="female"], na.rm = TRUE)

```

```{r playing_with_data}

cols = c(3,4,5,20:27)
sub2 = adults[ ,cols]
sub2 = scale(sub2)

sub2 = as.data.frame(sub2)


# this is the standard correlation matrix
sub2.corr = cor(sub2, use = "complete.obs");

#install.packages("KMO", dependencies=TRUE)
#library(KMO); # install.packages("KMO", dependencies=TRUE);  # not available for R == 4.0.2
#install.packages("REdaS", dependencies=TRUE);
library(REdaS); # install.packages("REdaS", dependencies=TRUE);

# https://www.rdocumentation.org/packages/REdaS/versions/0.9.3/topics/Kaiser-Meyer-Olkin-Statistics

sub2.KMO = KMOS(sub2, use = "complete.obs");


str(sub2.KMO);


my.kmo = sub2.KMO$KMO;
my.kmo;

if(my.kmo >= 0.90)
  {
  print("marvelous!");
  } else if(my.kmo >= 0.80)
    {
    print("meritorious!");
    }  else if(my.kmo >= 0.70)
        {
        print("middling!");
        } else if(my.kmo >= 0.60)
            {
            print("mediocre!");
            }  else if(my.kmo >= 0.50)
                {
                print("miserable!");
                } else { 
                        print("mayhem!");
                        print("Oh snap!"); 
print("Kaiser-Meyer-Olkin (KMO) Test is a measure of how suited your data is for Factor Analysis. The test measures sampling adequacy for each variable in the model and for the complete model. The statistic is a measure of the proportion of variance among variables that might be common variance. The lower the proportion, the more suited your data is to Factor Analysis. <https://www.statisticshowto.com/kaiser-meyer-olkin/>");
                }

```

```{r}
#review the correlations of various components of the body. 

#Correlation is a technique for investigating the relationship between two quantitative, continuous variables, for example, age and blood pressure. Pearson's correlation coefficient (r) is a measure of the strength of the association between the two variables.

#cor.test(x.0.1, y.0.1.max, method="pearson");  # default

#plot correlation
#corrplot::corrplot( (cor(iris.df)) ); 

```

```{r}
#Create a dendrogram/ evaluate clustering?
#dist(x)  Creates a distance function
#hclust creates a dendrogram.
#consider whether the data should be scaled scale(data)
#library(pvclust);  # install.packages("pvclust", dependencies=TRUE); #draws boxes around the clusters.  Gives fewer clusters than hclust
#pvclust uses "bootstrapping" to analyze the stability of the hclust.  For some reason, you have to transpose the data inputs.
#use ward.d2
```

```{r}
#Kmeans clustering is faster than hclust.  See notes at the end of instructor notebook for ideas.
```



You need to formulate some initial Research Questions that would interest you.  The power of "one" suggests there is an overall research question that may have sub-questions.

- e.g., Body Proportions and Artistic Renderings
- e.g., NBA Basketball Players vs Our Sample
- e.g., Studies of Covariates on various aspects of the body measurements

Notice the above are three example topics to be discovered, not an actual research question with sub-questions.  Given the data (and possibly external auxiliary data), what would you enjoy researching?  What would you find intriguing?



[+5 EASTER]  Using RGL or some other 3-D plotting tool, build a stick-figure based on your individual body measurements.  We assume the x-axis starts at the center of the eyes and moves laterally: looking right to the end of your outstretched fingertips is one direction; looking left to the end of your outstretched fingertips in another direction.  The y-axis would be the vertical axis from your feet to your head.  The z-axis would be forward-backward with zero being directly between your two feet (placed 'arm-pit' width apart), half of your shoe size being in the positive direction, the other half in the negative direction.  We could imaging z=0 running directly through your body (kinda like your spine and through your head).  Use line segments to connect parts, and a large point (or circle) where segments connect.  Use a rectangularish-shape for the head if a 3-D ellipsoid is too difficult).

https://en.wikipedia.org/wiki/Anatomical_terms_of_location

- x: lateral (right/left)
- y: superior/inferior
- z: anterior/posterior



